On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 tt_theorem.tex

\begin{theorem}[Tunnel Troll Theorem]
\label{TTT}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$. The following statements hold.
\begin{enumerate}
\item At arity $\alpha \geq 3$, if $S[i] = t$ (i.e. the tunnel that stabilizes $w[i]$ is not singular) and $S[i+1] \neq \blacksquare$, then $\#bc(C_{i-1}) > \#bc(C_i)$.
\item At arity $\alpha = 2$, if $n$ is the number of $bt$ in $S[i..j]$ for some $i,j$ such that $S[i] = b$ and $i < j$, then $\#bc(C_{i-2}) - \#bc(C_{j}) \geq n$.
\end{enumerate}
\end{theorem}

@@

\begin{lemma}
\label{TTT_tunnelC_lemma}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$, arity $\alpha = 2$.
If $S[i+2] = t_a$, the following statements hold.
\begin{enumerate}
\item If $w[i]$ forms at least one bond, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consumes any bond and $S[i] \in \{t_s, t_o \}$, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consumes any bond and $S[i] = t_a$, $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
\end{enumerate}
%If there are indices $i,j$ such that $S[i..j+1]$ is either $bbt^{(j-i-1)}b$ and $S[i+2] = t_a$ or $bt^\ell bt^{j-i-1-\ell}b$ for some $\ell$ and $S[i+l+2] = t_a$, then $\#bc(C_{i-1}) > \#bc(C_j)$.
\end{lemma}

@@

	We consider each statements. First we prove statement 1.
	The bead $w[i+1]$ consumes one hand and provide nothing.
	If $w[i]$ forms two bonds, then even if $w[i+2]$ provides two free hands, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	On the other hand, if it leaves a free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Next we prove statement 2.
	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $S$ or $O$.
	As prove in Lemma~\ref{TTT_exit}, $\#bc(C_{i-2}) \geq \#bc(C_{i})$.%the increase of binding capability by $w[i-1]$ and $w[i]$ is at most 0.
	The bead $w[i]$ leaves two free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Finally, we prove statement 3.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $A$ and $w[i]$ does not consumes any bond.
	Then, $w[i-2]$ must consume two bonds; otherwise $w[i-2]$'s hands is used by $w[i]$.
	The bead $w[i-1]$ consumes one hand and provide nothing.
	Thus,  $\#bc(C_{i-3}) - 1 \geq \#bc(C_{i})$.
	Moreover, the bead $w[i+1]$ consumes one hand and provide nothing and $w[i]$'s hands is used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Therefore, $\#bc(C_{i-3}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.
\end{proof}

\begin{figure}[tb]
  \begin{center}
        \begin{tikzpicture}
        \draw[->,white] (-70:0.1) -- (-70:3);
          \foreach \theta in {60,-60,120,0}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

@@

          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];
            \foreach \theta in {-120,-60,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[i+2] = t_a$.}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

Next, we consider the case of $\delta = 2, \alpha = 1$. We assume there are an indices $i$ ,$j$ such that $S[i+1..j+1] = bt^{(j-i-1)}b$ and all $S[i+2..j]$ are  $ t_s$ or  $t_o$.
According to Lemma~\ref{TTT_entrance_Tab}, $\#bc(C_{i-1}) > \#bc(C_{i})$ and moreover that, according to Lemma~\ref{TTT_exit}, $\#bc(C_{i}) \geq \#bc(C_{j})$.
Thus, binding capability of the conformation decrease per a subsequence $bt_s$ or $bt_o$.

On the other hand, we assume there are an indices $i$ ,$j$ such that $S[i+1..j+1] = bt^{(j-i-1)}b$ and there is a element of $S[i+2..j]$ such that it is $t_a$.
All $S[i+3..j]$ do not contain $t_o$; otherwise $S[i+4..j+1]$ contain $\blacksquare$.
Therefore, $S[i+2] = t_a$.
According to Lemma~\ref{TTT_exit}, on statement 1, 2, $\#bc(C_{i-2}) > \#bc(C_{i+2})$ and if $i \leq j - 3$, $\#bc(C_{i+2}) \geq \#bc(C_{j})$.
Thus, binding capability of the conformation decrease per a subsequence $bt_a$ unless $w[i]$ does not consumes any bond.
If $w[i]$ does not consumes any bond and $S[i] = t_a$, that is, $S[i-1..i+2] = bt_abt_a$, then $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
In this case, two subsequences of $bt_a$ collaborate decreasing binding capability but eventually, two subsequences decrease two bond.
That is, one subsequence decrease binding capability.

Hence, if $S[i..j]$ contain n's subsequences of $bt$, $\#bc(C_{i-2}) - \#bc(C_{j}) \geq n$
\qed

>>>>

\begin{theorem}[Tunnel Troll Theorem]
\label{TTT}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$. The following statements hold.
\begin{enumerate}
\item At arity $\alpha \geq 3$, if $S[i] = t$ (i.e. the tunnel that stabilizes $w[i]$ is not singular) and $S[i+1] \neq \blacksquare$, then $\#bc(C_{i-1}) > \#bc(C_i)$.
\item At arity $\alpha = 2$, if $m$ is the number of occurrences of $bt$ as a factor in $S[1..k]$ for an index $k$, then $\#bc(C_{0}) - m \geq \#bc(C_{k})$.
\end{enumerate}
\end{theorem}

In order to prove this Tunnel Troll Theorem, we use the following three lemmas.
%Let us introduce some lemmas for proving Tunnel Troll Theorem.

@@

\begin{lemma}
\label{TTT_tunnelC_lemma}
Let $\Xi$ be a deterministic unary oritatami system of delay $\delta = 1$, arity $\alpha = 2$.
If $S[i+2] = t_a$, the following statements hold.
\begin{enumerate}
\item If $w[i]$ forms at least one bond, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consume any bond and $S[i] \in \{t_s, t_o \}$, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
\item If $w[i]$ does not consume any bond and $S[i] = t_a$, $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
\end{enumerate}
%If there are indices $i,j$ such that $S[i..j+1]$ is either $bbt^{(j-i-1)}b$ and $S[i+2] = t_a$ or $bt^\ell bt^{j-i-1-\ell}b$ for some $\ell$ and $S[i+l+2] = t_a$, then $\#bc(C_{i-1}) > \#bc(C_j)$.
\end{lemma}

@@

	We consider each statement. First we prove Statement 1.
	The bead $w[i+1]$ consumes one hand and provides nothing.
	If $w[i]$ forms two bonds, then even if $w[i+2]$ provides two free hands, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	On the other hand, if it leaves a free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-1}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Let us proceed to Statement 2.
	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $S$ or $O$.
	As prove in Lemma~\ref{TTT_exit}, $\#bc(C_{i-2}) \geq \#bc(C_{i})$.%the increase of binding capability by $w[i-1]$ and $w[i]$ is at most 0.
	The bead $w[i]$ leaves two free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Finally, we prove Statement 3.
	%Consider the case when $w[i]$ is stabilized by a tunnel section of type $A$ and $w[i]$ does not consumes any bond.
	In order for $w[i]$ not to bind, $w[i-2]$ must have already used up its hands.
	% The bead $w[i-2]$ must consume two bonds ; otherwise $w[i-2]$'s hands is used by $w[i]$.
	The bead $w[i-1]$ consumes one hand and provides nothing.
	Thus,  $\#bc(C_{i-3}) - 1 \geq \#bc(C_{i})$.
	The bead $w[i+1]$ consumes one hand and provides nothing.
	Finally, $w[i+2]$ uses one hand of $w[i]$, and hence, does not increase the binding capability.
	Therefore, $\#bc(C_{i-3}) -2 \geq \#bc(C_{i+2})$.
	This argument actually works also for the case when $w[i]$ is stabilized rather by binding. \qed
\end{proof}

\begin{figure}[tb]
  \begin{center}
  \begin{tabular} {c c}
  \begin{minipage}{0.48\hsize}
  \centering
        \begin{tikzpicture}
        \draw[->, white, shift=(180:1), shift=(120:1)] (120:0.9) -- (120:0.1);
          \foreach \theta in {60,-60,120,0}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

@@

          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];

          \end{scope}
        \end{tikzpicture}

	\end{minipage}

	\begin{minipage}{0.48\hsize}     
        \begin{tikzpicture}
         \draw[->, white, shift=(180:1), shift=(120:1)] (120:0.9) -- (120:0.1);
          \foreach \theta in {60,-60,120,0}{
            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
          }

          \foreach \theta in {60,60,120,180}{
            \fill[transform canvas={shift=(\theta:1)}, shift=(180:1)](120:1) circle [radius=0.1];
          }

          \fill[blue](180:1) circle [radius=0.05];
          \fill[blue](0:0) circle [radius=0.05];
          \fill[blue, shift=(180:1), shift=(120:1)](0:0) circle [radius=0.05];
          \fill[blue, shift=(180:1), shift=(120:1)](-120:1) circle [radius=0.05];

          \draw[->, blue] (180:0.9) -- (180:0.1);
          \draw[->, blue] (-120:0.1) -- (-120:0.9);
          \draw[->, blue, shift=(180:1), shift=(120:1)] (-60:0.1) -- (-60:0.9);
           \draw[->, blue, shift=(180:1), shift=(120:1)] (-120:0.9) -- (-120:0.1);

          \node[below] at (180:1) {$w[i]$};
          \node[above] at (0:0) {$w[i+1]$};
          \node[below] at (-120:1) {$w[i+2]$};
          \node[below left] at (-180:2) {$w[i-2]$};

          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];

          \end{scope}
        \end{tikzpicture}
        \end{minipage}
        \end{tabular}
    \caption{Case of $S[i] = t_a$ and $S[i+2] = t_a$: (Right) Moreover $w[i]$ does not consume any bond.}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

Next, we consider the case of $\delta = 2, \alpha = 1$. We assume there are indices $i$, $j$ such that $S[i+1..j+1] = bt^{(j-i-1)}b$.
If $S[i+2]$ is $ t_s$ or  $t_o$, then Lemma~\ref{TTT_entrance_Tab} implies $\#bc(C_{i-1}) > \#bc(C_{i})$ and  Lemma~\ref{TTT_exit} implies $\#bc(C_{i}) \geq \#bc(C_{j})$.
Thus, binding capability decreases by 1 per a factor $bt_s$ or $bt_o$.

Now, we assume $S[i+2] = t_a$.
%All $S[i+3..j]$ do not contain $t_a$; otherwise $S[i+4..j+1]$ contain $\blacksquare$.
%Therefore, $S[i+2] = t_a$.
According to Lemma~\ref{TTT_tunnelC_lemma}, on statement 1, 2, $\#bc(C_{i-2}) > \#bc(C_{i+2})$ and if $i \leq j - 3$, $\#bc(C_{i+2}) \geq \#bc(C_{j})$.
Thus, binding capability of the conformation decrease per a subsequence $bt_a$ unless $w[i]$ does not consumes any bond.
If $w[i]$ does not consumes any bond and $S[i] = t_a$, that is, $S[i-1..i+2] = bt_abt_a$, then $\#bc(C_{i-3}) - 2 \geq \#bc(C_{i+2})$.
In this case, two subsequences of $bt_a$ collaborate decreasing binding capability but eventually, two subsequences decrease two bond.
That is, one subsequence decrease binding capability.

Hence, if $S[i..j]$ contain n's subsequences of $bt$, $\#bc(C_{i-2}) - n \geq \#bc(C_{j})$
\qed
